---
title: "SNP Quality Control Filtering - Grevillea aquifolium"
author: "Simon E. Lobos"
date: "June, 20267"
output: html_document
---

Import packages
```{r include = TRUE, message=FALSE, eval = FALSE}
library(dartR)
library(ggplot2)
library(adegenet)
library(ape)
```

Import data
```{r}
#Set working directory 
setwd("~/Desktop/Comparative-Landscape-Genomics/1_SNP_quality_control_filtering/GA")  

#Import the DArT genotype CSV file and the complementary CSV file containing sample metadata.
gl <- gl.read.dart(filename = "Report_DGr22-6726_SNP_2.csv", ind.metafile = "GA_metadata.csv")

#Print details
gl

#Check the order of individual names. If individuals are not sorted alphabetically, proceed to the next step.
indNames(gl)

#Sort all individuals by population in alphabetical order
gl <- gl[order(pop(gl)), ]

#Verify the new ordering of individuals
indNames(gl)
```

Check for potential clone mates
```{r }
#Method 1 - calculating the proportion of shared alleles
clones <- gl.propShared(gl)

#Convert matrix into data frame 
clones_temp <- as.data.frame(as.table(as.matrix(clones)))
write.csv(clones_temp, file = "GA_clones.csv")

#Method 2 - Neighbour-joining tree
NJ <- dist(tab(gl))
tree <- nj(NJ)
plot(tree, cex=0.5)

#Export tree into Nexus format for visualisation in FigTree
write.nexus(trec, file = "GA_clones_NJ.nex") 
```

Remove potential clones
```{r}
gla <- gl.drop.ind(gl, ind.list=c("GA_8_4B","GA_38A_5A","GA_42_1B",
                                  "GA_76A_6A","GA_84_4A","GA_95_6B",
                                  "GA_101B_2A","GA_107B_1B")) 

#Print details
gla
```

Reports on minor allele frequencies (maf) overall, and by population
```{r }
pdf("GA_maf_gla.pdf", height=20, width=20)
gl.report.maf(gla)
dev.off()
```

Filters loci by maf
```{r }
pdf("GA_maf_glb.pdf")
glb <- gl.filter.maf(gla, threshold = 0.02)
dev.off()

#Print details
glb
```

Report Call Rates/ Missing Data - loci
```{r }
pdf("GA_report_callrate_loc_glb.pdf")
gl.report.callrate(glb, method='loc')
dev.off()
```

Report Call Rates/ Missing Data - individuals
```{r }
pdf("GA_report_callrate_ind_glb.pdf")
gl.report.callrate(glb, method='ind')
dev.off()
```

Calculate call rate for each locus
```{r }
pdf("GA_callrate_loc_glc.pdf")
glc <- gl.filter.callrate(glb, method="loc", threshold=0.8)
dev.off()

#Print details
glc
```

Calculate call rate for each individual
```{r }
pdf("GA_callrate_ind_gld.pdf")
gld <- gl.filter.callrate(glc, method="ind", threshold=0.8)
dev.off()

#Print details
gld
```

Discard loci that might be possible paralogues (hamming distance)
```{r }
pdf("GA_paralogues_gle.pdf")
gle <- gl.filter.hamming(gld, threshold=0.2)
dev.off()
 
#Print details
gle

save(gle, file="GA_gle.rdata")
```

Compare smear plots between the original and filtered datasets
```{r }
pdf("GA_original_dataset_gl.pdf")
plot(gl) 
dev.off() 

pdf("GA_filtered_dataset_gle.pdf")
plot(gle)
dev.off()
```


