---
title: "SNP Quality Control Filtering - Eucalyptus obliqua"
author: "Simon E. Lobos"
date: "June, 20267"
output: html_document
---

Import packages
```{r include = TRUE, message=FALSE, eval = FALSE}
library(dartR)
library(ggplot2)
library(adegenet)
library(ape)
```

Import data
```{r}
#Set working directory 
setwd("~/Desktop/Comparative-Landscape-Genomics/1_SNP_quality_control_filtering/EO")  

#Import the DArT genotype CSV file and the complementary CSV file containing sample metadata.
gl <- gl.read.dart(filename = "Report_DBank22-6724_SNP_2.csv", ind.metafile = "EO_metadata.csv")

#Print details
gl

#Check the order of individual names. If individuals are not sorted alphabetically, proceed to the next step.
indNames(gl)

#Sort all individuals by population in alphabetical order
gl <- gl[order(pop(gl)), ]

#Verify the new ordering of individuals
indNames(gl)
```

Check for potential clone mates
```{r }
#Method 1 - calculating the proportion of shared alleles
clones <- gl.propShared(gl)

#Convert matrix into data frame 
clones_temp <- as.data.frame(as.table(as.matrix(clones)))
write.csv(clones_temp, file = "EO_clones.csv")

#Method 2 - Neighbour-joining tree
NJ <- dist(tab(gl))
tree <- nj(NJ)
plot(tree, cex=0.5)

#Export tree into Nexus format for visualisation in FigTree
write.nexus(trec, file = "EO_clones_NJ.nex") 
```

Remove potential clones
```{r}


gla <- gl.drop.ind(gl, ind.list=c("EO_8_2B","EO_27A_1A","EO_27A_1B","EO_52_6B","EO_76A_3B",
                                  "EO_84_3B","EO_98B_1A","EO_104_5A","EO_106A_4A","EO_87A_1",
                                  "EO_87A_6","EO_93A_1","EO_93A_2","EO_93A_3","EO_93A_4",
                                  "EO_93A_5","EO_93A_6","EO_105C_1","EO_105C_2","EO_105C_3",
                                  "EO_105C_4","EO_105C_5","EO_105C_6","EO_27A_5","EO_27A_6",
                                  "EO_27A_2","EO_27A_3","EO_27A_4","EO_98B_5","EO_45_5",
                                  "EO_4_1","EO_4_2","EO_4_6","EO_4_5","EO_4_3",
                                  "EO_40A_2","EO_40A_4","EO_40A_5","EO_40A_6","EO_24A_4",
                                  "EO_40A_1","EO_24A_1","EO_24A_5","EO_24A_3","EO_100A_1",
                                  "EO_100A_3","EO_104_4","EO_45_1","EO_45_3","EO_95_6",
                                  "EO_AOS03B_2","EO_8_5","EO_15_5","EO_49AB_4","EO_69A_5",
                                  "EO_98B_3")) 

#Print details
gla
```

Reports on minor allele frequencies (maf) overall, and by population
```{r }
pdf("EO_maf_gla.pdf", height=20, width=20)
gl.report.maf(gla)
dev.off()
```

Filters loci by maf
```{r }
pdf("EO_maf_glb.pdf")
glb <- gl.filter.maf(gla, threshold = 0.02)
dev.off()

#Print details
glb
```

Report Call Rates/ Missing Data - loci
```{r }
pdf("EO_report_callrate_loc_glb.pdf")
gl.report.callrate(glb, method='loc')
dev.off()
```

Report Call Rates/ Missing Data - individuals
```{r }
pdf("EO_report_callrate_ind_glb.pdf")
gl.report.callrate(glb, method='ind')
dev.off()
```

Calculate call rate for each locus
```{r }
pdf("EO_callrate_loc_glc.pdf")
glc <- gl.filter.callrate(glb, method="loc", threshold=0.8)
dev.off()

#Print details
glc
```

Calculate call rate for each individual
```{r }
pdf("EO_callrate_ind_gld.pdf")
gld <- gl.filter.callrate(glc, method="ind", threshold=0.8)
dev.off()

#Print details
gld
```

Discard loci that might be possible paralogues (hamming distance)
```{r }
pdf("EO_paralogues_gle.pdf")
gle <- gl.filter.hamming(gld, threshold=0.2)
dev.off()
 
#Print details
gle

save(gle, file="EO_gle.rdata")
```

Compare smear plots between the original and filtered datasets
```{r }
pdf("EO_original_dataset_gl.pdf")
plot(gl) 
dev.off() 

pdf("EO_filtered_dataset_gle.pdf")
plot(gle)
dev.off()
```


