---
title: "LFMM2 - Callitris rhomboidea"
author: "Simon E. Lobos"
date: "June, 2026"
output: html_document
---

Code adapted from Ahrens et al. (2021).https://doi.org/10.1111/1755-0998.13351

Import packages
```{r include = TRUE, message=FALSE, eval = FALSE}
library(dartR) 
library(LEA)
library(lfmm)
library(dplyr)
library(tibble)
library(stringr)
library(ggplot2)
```

Import data
```{r }
#Set directory
setwd("~/Desktop/Comparative-Landscape-Genomics/3_GEA/3_2_LFMM/CR")

#Load genlight object
load("~/Desktop/Comparative-Landscape-Genomics/1_SNP_quality_control_filtering/CR/CR_gle.rdata")

#Convert genlight to geno object
CR.geno<-gl2geno(CR_gle, outpath = getwd()) 
```

SNMF - Inference of individual admixture coefficients
```{r }
project = NULL
project = snmf("gl_geno.lfmm", K=1:30, ploidy = 2, entropy = T, alpha=100, rep=100, CPU=4, project = "new")

#To load the project, use:
project = load.snmfProject("gl_geno.snmfProject")
```

Identify ancestral populations (K) 
```{r }
#Plot cross-entropy criterion for each K
K <- summary(project)$crossEntropy %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("temp") %>%
  mutate(K = as.numeric(str_extract(temp, "(\\d)+"))) %>%
  select(-temp)

#Choose K for which the function plateaus or increases sharply
ggplot(K, aes(x = K, y = mean)) +
  geom_line(color = "black", size = 0.25 ) +
  geom_segment(aes(x = K, y = min, xend = K, yend = max)) +
  geom_point(shape = 21, size = 4, color = "black", fill = "white") +
  geom_point(data=K[1, ], aes(x = K, y = mean), shape = 21, size = 4, color = "black", fill = "#44ba8e") +
  scale_x_continuous(breaks = seq(1, 30, by = 1)) +
    ylim(0.45, 0.80) +
  xlim(1, 30) +
  #labs(x = "Number of ancestral populations", y = "Cross-entropy") +
  xlab("") +
  ylab("") +
  theme_classic() + 
  theme(legend.position="none",
        legend.title = element_text(size = 18),
        axis.text.x = element_text(colour="black", size=18, vjust = 0.5, hjust = 0.5),
        axis.text.y = element_text(colour="black", size=18, vjust = 0.5, hjust = 0.5),
        axis.title.y = element_text(colour="black", size=18, face="plain"),
        axis.title.x = element_text(colour="black",size=18,face="plain"),
        plot.title = element_text(color="black", size=20, face="italic")) +
  ggtitle("Callitris rhomboidea")

ggsave("LEA_cross_entropy_CR.pdf")
```

Select the best K with the lowest cross-entropy value
```{r }
best = which.min(cross.entropy(project, K = 1))
```

Impute the missing genotypes using snmf project previously created
```{r }
impute(project, "gl_geno.lfmm",
       method = 'mode', K = 1, run = best)
```

Proportion of correct imputation results
```{r }
x = read.table("gl_geno.lfmm_imputed.lfmm", header=F)
sum(is.na(x)) # 0
x[1:10,1:10]
```

Run LFMM2
```{r }
#Import environmental population data (no row/column names)
E<-read.csv("~/Desktop/Comparative-Landscape-Genomics/2_Environmental_data/CR_env_lfmm.csv", header=F)

head(E)

#Create directories
#calpval: list of SNPs and show the ones that are potential adaptive 
dir.create('./calpval/')
dir.create('./prs/')

for (i in 1:length(colnames(E))) { #4 each env variable
  #### Run LFMM2 #####
  mod.lfmm <- lfmm_ridge(Y = x, 
                         X = E[[i]],
                         K = 1) #chose K based on PCA
  ## performs association testing using the fitted model:
  pv <- lfmm_test(Y = x, #x=genotype
                  X = E[[i]], #y=env
                  lfmm = mod.lfmm, 
                  calibrate = "gif")
  ## record the calibrated pvalues
  pvalues <- pv$calibrated.pvalue 
  hist(pvalues)
  length(which(pvalues < 0.05))
  write.table(pvalues, paste0("./calpval/cal_",colnames(E)[i],".calpval"), quote = FALSE)
  #polygenic risk scores
  x.pred <- predict_lfmm(x,E[[i]], mod.lfmm, fdr.level = 0.05, newdata = NULL) #similar to R2
  
  #perform lm on the predicted values vs env (i.e. can snps predict environment)
  x.lm <- lm(x.pred$pred ~ scale(E[[i]], scale = FALSE))
  #calculate r2 for the polygenic risk score
  prs <- summary(x.lm)$adj.r.squared #Different to RDA 
  write.table(prs, paste0("./prs/cal_",colnames(E)[i],".prs"), quote = FALSE)
}
```

Combine p-values in a single
```{r }
file.list<- list.files("calpval/", pattern=".calpval")

test_read <- read.table(file=paste0("calpval/",file.list[1]), header=T)

#get the bios
library(stringr)
fl <- word(file.list, 2, sep = "_")
fl2 <- gsub('.calpval', '', fl)

#read tables in
dataMerge <- data.frame(name=paste0("V",1:dim(test_read)[1]))

for(f in 1:length(file.list)){ 
  ReadInMerge <- read.table(file=paste0("calpval/",file.list[f]), header=T)
  colnames(ReadInMerge)[1] <- fl2[f]
  dataMerge <- cbind.data.frame(dataMerge,ReadInMerge)
}

all.pv <- dataMerge

write.csv(all.pv, "cal_all_pv.csv", quote=FALSE)

all.pv <- read.csv("cal_all_pv.csv") #Check snps are ordered
all.pv <- all.pv[,-1:-2]
head(all.pv)

colnames(all.pv) <- c("TWI","AVP","pHc","BIO01","BIO05","BIO12","BIO14")

#include name of the snps
ind_snp <- read.csv("~/Desktop/Comparative-Landscape-Genomics/3_GEA/3_1_BayPass/CR/CR_loc.names.csv")[-1]

#rename the column
colnames(ind_snp) <- "ind_snp"

row.names(all.pv) <- ind_snp[,1]

write.csv(all.pv, "./cal_all_pval.csv", quote=FALSE)

all.pv <- data.frame(all.pv) 
```

Time to pick the variables to use
```{r }
all_names <- colnames(all.pv)
sel_bio <- all_names[1:4] 
v0 <- NULL

for (i in sel_bio) {
  v0 <- c(v0, which(all.pv[,i] < 0.001))
  cat(length(v0), "\n")
}

uniq_snps <- unique(v0); length(uniq_snps) 
x = read.geno("gl_geno.geno")

x_t <- data.frame(t(x)); dim(x_t)

sig_df_1 <- x_t[uniq_snps,]; dim(sig_df_1) 

write.csv(sig_df_1, file=paste0("CR_outliers_lfmm2.csv"))

sig_df_1[1:10,1:10]
```

Check SNPs under putatevely selection by environmental variable
```{r }
nrow(subset(all.pv, TWI < 0.001)) 
nrow(subset(all.pv, AVP < 0.001)) 
nrow(subset(all.pv, pHc < 0.001)) 
nrow(subset(all.pv, BIO05 < 0.001)) 

out.TWI <- (subset(all.pv, TWI < 0.001)) 
out.AVP <- (subset(all.pv, AVP < 0.001)) 
out.pHc <- (subset(all.pv, pHc < 0.001)) 
out.BIO05 <- (subset(all.pv, BIO05 < 0.001)) 

out.TWI.df <- data.frame(rownames(out.TWI))
out.AVP.df <- data.frame(rownames(out.AVP))
out.pHc.df <- data.frame(rownames(out.pHc))
out.BIO05.df <- data.frame(rownames(out.BIO05))

data_names <- c("out.TWI.df","out.AVP.df","out.pHc.df","out.BIO05.df")

for(i in 1:length(data_names)) {                              # Head of for-loop
  write.csv2(get(data_names[i]),                              # Write CSV files to folder
             paste0(data_names[i],
                    ".csv"),
             row.names = FALSE)
}
```

